Bootstrapping Multi-atlas Segmentation Using Multiple Automatically Generated Templates for the Segmentation of the Whole Hippocampus and Subfields

### 2013-12-23 paper.Rnw
subfield segmentation plots

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% ADNI1 X-Validation Dataset Demographics Table
%
<<ADNI-xval-demographics, echo=F, dependson='ADNI1-xval-load', results="asis",cache=FALSE>>=
tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + CDRSB + ADAS13 +
               MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
               jens_atlases$RID, method = "reverse", test=FALSE, overall=TRUE)
caption = 
  "\\textbf{ADNI-1 cross-validation subset demographics.} 
  CN     - Cognitively Normal. 
  LMCI   - Late-onset Mild Cognitive Impairment. 
  AD     - Alzheimer's Disease.  
  Hisp   - Hispanic. 
  CDR-SB - Clinical Dementia Rating-Sum of Boxes.  
  ADAS   - Alzheimer's Disease Assessment Scale.  
  MMSE   - Mini-Mental State Examination."
latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
      caption=caption, caption.loc = c("top"),
      label="tab:ADNI1-xvalidation-demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Subfield Validation Template library demographics (healthy controls)
%                                     
<<WAval-healthy-demographics, echo=F, results="asis",cache=FALSE>>=
master = read.csv('data/DT_study.csv')
tab <- summary( Diagnosis ~ Age + Sex + Education + Handedness, data = master,
    method="reverse", na.rm = TRUE, test=FALSE)
caption = 
  "\\textbf{Demographics for the hippocampal subfield cross-validation healthy
control subject sample used in the template library (excluding the Winterburn
atlas subjects).} Education is shown in years."
latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
      caption=caption, caption.loc = c("top"),
      title="tab:WAval-healthy-demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First Episode SZ Demographics Table
%
<<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
caption = 
  "\\textbf{Schizophrenia First Episode Patient Demographics.}
  ambi - ambidextrous. 
  SES  - Socioeconomic Status score. 
  FSIQ - Full Scale IQ.
"
fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
fep_sz_demographics = subset(fep_sz_demographics, Anon <= 81)
tab <- summary(DX ~ Age + Gender + Handedness + Education + SES + FSIQ,
               data=fep_sz_demographics, method = "reverse", test = FALSE)
latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
      caption=caption, caption.loc = c("top"),
      title="tab:SZFEP-Demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                             
% ADNI1 Complete 1Yr Dataset Demographics Table 
%                                                                             
<<ADNI1-Dataset-Demographics, echo=F, results="asis",cache=FALSE>>=
# RID < 2000 ensures only ADNI1 patients
yr1 = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
yr1$VISCODE <- factor(yr1$Visit, levels = c(1,2,3,4), labels=c("bl","m03","m06","m12"))
yr1 = subset(yr1, !grepl("Scaled_2", Description))
yr1 = yr1[,c("RID","VISCODE")]
yr1 = 
adnimerge.yr1 = merge(adnimerge, yr1) 
tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + CDRSB + ADAS13 +
               MMSE, data = adnimerge.yr1, 
               #subset = RID %in% yr1$RID & VISCODE %in% c('bl','m06','m12'),   
               method = "reverse", test=FALSE, overall=TRUE)
caption = 
  "\\textbf{ADNI1 1.5T Complete 1Yr dataset demographics.} 
  CN     - Cognitively Normal. 
  LMCI   - Late-onset Mild Cognitive Impairment. 
  AD     - Alzheimer's Disease.  
  Hisp   - Hispanic. 
  CDR-SB - Clinical Dementia Rating-Sum of Boxes.  
  ADAS   - Alzheimer's Disease Assessment Scale.  
  MMSE   - Mini-Mental State Examination."
latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
    caption=caption, caption.loc = c("top"),
    title="tab:ADNI1-Dataset-Demographics",
    label="tab:ADNI1-Dataset-Demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: show variability of MAGeT over parameters
\begin{figure}
<<ADN1-Xval-variability,cache=TRUE,dependson='ADNI1-Xval-prep',fig.width=5,fig.height=5>>=
ants = subset(jens_xval_data, 
                     reg_method=="ANTS" & 
                     method.mb=="Majority Vote" &
                    templates.mb %in% seq(1,20,2))

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%                                                                     %%%%%
%%%%%    RESULTS: Experiment 2: Hippocampal Subfield Cross-Validation     %%%%%
%%%%%                                                                     %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment 2 Results: Winterburn Atlases Cross-Validation}
<<WAval-prep, cache=T>>=
wa.xval.vols = subset(read.csv('data/WAval.csv'))
wa.xval.kappa = subset(read.csv('data/WAval-kappa.csv'))
wa.xval.casted = dcast(melt(wa.xval.vols, id.vars=c("file","version")), file + variable ~ version)
wa.xval.casted = rename(wa.xval.casted, c("variable"="region"))
levels(wa.xval.casted$region) <-list("CA1"="X1", "CA1"="X101", 
                            "Subiculum"="X2","Subiculum"="X102", 
                            "CA4/DG"="X4","CA4/DG"="X104", 
                            "CA2/CA3"="X5","CA2/CA3"="X105", 
                            "SR/SL/SM"="X6","SR/SL/SM"="X106")
@
This experiment explores \mb{} segmentation of hippocampal subfields. To achieve
this, a leave-one-out validation was conducted in which lower-resolution images
($0.9mm^3$ voxels) of each Winterburn atlas subject were segmented using the
remaining high-resolution Winterburn atlas subjects' images.

In general, across hippocampal subfields the percent error in volume of \mb{}
segmentations (relative to the full-resolution Winterburn atlas segmentation)
compares favourably to the error resulting from image resampling (Figure
\ref{fig:WAval-vol-boxplot}). In particular, there is no significant difference 
between percent error of \mb{} volumes compared to image resampling error for the 
CA1, Subiculum, and CA2/CA3 subregions. The \mb{} volumes of CA4/DG and SR/SL/SM 
subregions show both significantly different (lower) percent error (p < 0.001).

Figure \ref{fig:subfield-montage} shows a qualitative comparison of \mb{} subfield
segmentations for a single subject.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Winterburn-XVAL: compare resampling
\begin{figure}[h!]
<<WAval-vol-boxplot,cache=T,fig.width=6,fig.height=5>>=
wa.xval.error = ddply(wa.xval.casted, c("file", "region"), function(df) {
  with(df, 
  data.frame(
    resampled = (gold_0.9mm - gold_0.3mm) / gold_0.3mm * 100, 
    mb_bravo  = (mb_bravo   - gold_0.3mm) / gold_0.3mm * 100, 
    mb_0.9mm  = (mb_0.9mm   - gold_0.3mm) / gold_0.3mm * 100))})

@@

ggplot(wa.xval.error.melted, aes(y=percenterr,x=measure,colour=measure)) +
  facet_grid(.~region) + 
  geom_hline(aes(yintercept=0), colour="grey", size=1) + 
  geom_boxplot() + 
  geom_point(group=1, position="dodge") + 
  labs(y="Unilateral subregion volume percent error", x="Subfield") +
  theme(axis.ticks = element_blank(), axis.text.x = element_blank()) + 
  scale_colour_discrete("") + 
  theme(legend.direction = "vertical", legend.position = "bottom")
@
  \caption{{\bf Percent error in unilateral subfield volume.} 
  Percent error is measured against the volumes of the unmodified Winterburn
  atlas subfield segmentations.
  {\bf Subsampled atlas labels} volumes of the manual segmentations of the Winterburn
  atlases after resampling to $0.9mm^3$ voxels.  
  {\bf MAGeT on subsampled atlas images} volumes are \mb{} segmentations of the Winterburn atlas
  images after resampling to $0.9mm^3$ voxels.
  {\bf MAGeT on T1 BRAVO} volumes are \mb{} segmentations of T1 BRAVO images
  ($0.9mm^3$ voxels) acquired separately of four of the five Winterburn atlas
  subjects.} 
  \label{fig:WAval-vol-boxplot}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Winterburn-XVAL: volume plots
\begin{figure}[h!]
<<WAval-vol-plots,cache=T,fig.width=7,fig.height=6>>=
wa.xval.volumes = melt(wa.xval.casted, id.vars=c('file', 'region', 'gold_0.3mm'))
wa.xval.volumes = rename(wa.xval.volumes, c("variable"="measure", "value"="volume"))
levels(wa.xval.volumes$measure) <- list("Subsampled atlas labels"="gold_0.9mm",
                               "MAGeT on subsampled atlas images" = "mb_0.9mm",
                               "MAGeT on T1 BRAVO images"="mb_bravo")
wa.xval.vols.lm = ddply(wa.xval.volumes, c("region", "measure"), function(df) {
   model = lm(volume ~ gold_0.3mm, df)
   pearson = cor(df$volume, df$gold_0.3mm, use='pairwise.complete.obs')
   data.frame(
     eqn  = lm_eqn(model), 
     corr = pearson)})
wa.xval.vols.lm$yoffset = c( 1225 - 0:2 * 40,
                             465  - 0:2 * 15, 
                             750  - 0:2 * 30,
                             325  - 0:2 * 18, 
                             1100 - 0:2 * 40 )
wa.xval.vols.lm$xoffset = c( rep(700, 3),
                             rep(250, 3),
                             rep(490, 3),
                             rep(140, 3), 
                             rep(570, 3) )
ggplot(wa.xval.volumes, aes(x=gold_0.3mm, y=volume, colour=measure)) + 
  facet_wrap(~region, ncol=3, scales='free') + 
  geom_point() + 
  coord_equal() + 
  geom_smooth(method='lm')  + 
  xlab(expression("Manual Unilateral Volume" (mm^3))) +
  ylab(expression("Automated Unilateral Volume" (mm^3))) +
  geom_text(aes(x=xoffset, y=yoffset, label=eqn), parse=TRUE,
    data=wa.xval.vols.lm, size=2,hjust=0) +
  scale_color_discrete(name="") + 
  theme(legend.direction = "horizontal", legend.position = "bottom")

@
  \label{fig:WAval-volumes}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Winterburn-XVAL: bland-altman 
\begin{figure}[h!]
<<WAval-bland,cache=T,fig.width=7,fig.height=6>>=
wa.xval.volumes = melt(wa.xval.casted, id.vars=c('file', 'region', 'gold_0.3mm', 'gold_0.9mm'))
wa.xval.volumes = rename(wa.xval.volumes, c("variable"="measure", "value"="volume"))
levels(wa.xval.volumes$measure) <- list("Subsampled atlas labels"="gold_0.9mm",
                               "MAGeT on subsampled atlas images" = "mb_0.9mm",
                               "MAGeT on T1 BRAVO images"="mb_bravo")

@@

ggplot(wa.xval.volumes, aes(x=mean, y=diff, colour=measure)) + 
  facet_wrap(~region, ncol=3, scales='free') + 
  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
  geom_point(size=1) +  
  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
             data=wa.xval.bland.limits, size=0.3, alpha=0.5) + 
  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", 
             data = wa.xval.bland.limits, hjust=1.1,vjust=-0.8, size=2.5) + 
  xlab(expression('Mean of manual and automated HC volume ' (mm^3))) + 
  ylab(expression('Difference of manaul and automated HC volume ' (mm^3))) + 
  scale_color_discrete(name="") + 
  theme(legend.direction = "horizontal", legend.position = "bottom", 
        axis.text.y = element_text(angle = 90, hjust=0.5), 
        axis.text = element_text(size = 8))

@
  \label{fig:WAval-bland}
\end{figure}

@@

\begin{figure}
<<FEP-volumes,cache=FALSE,dependson="setup",fig.width=5,fig.height=5>>=
SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")
mean_sz_vols = ddply(SZ_volumes, c('Subject', 'Method'), function (df) {
  data.frame(Volume = mean(df$Volume))})

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1 1yr Complete Dataset Package Totals & Failures
%                        
<<ADNI-seg-package-totals, echo=F, results="asis",cache=FALSE>>=
package_totals$Total <- NULL
latex(package_totals, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
    caption="\\textbf{Number of segmented images and quality control failures of 
             \\adnidataset{} dataset by method.}", 
    caption.loc = c("top"), rowname=NULL, 
    label="tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h]
<<ADNI-volumes-boxplot,dependson='ADNI-volumes-prep',cache=T,fig.width=7,fig.height=4>>=
melted = melt(means.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT"),
              variable.name="Method", value.name="Volume")
correlations = ddply(melted, c("DX","Method"), function (df) {
  data.frame(
    pearson = cor(df$SNT, df$Volume)
  )
})
m2 = melt(means.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT","SNT"), variable.name="Method")
levels(m2$Method) = c("FS"    = "FreeSurfer", 
                      "FSL"   = "FSL", 
                      "MAPER" = "MAPER", 
                      "MAGeT" = "MAGeT", 
                      "SNT"   = "SNT")
m2$DX = factor(m2$DX, levels(m2$DX)[c(2,3,1)])

@@

\begin{figure}[h]
<<ADNI-volumes-plot,dependson='ADNI-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
melted=melt(means.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
            variable.name = "Method", value.name = "volume")

@@

\begin{figure}
<<ADNI-Bland-Altman, cache=TRUE, dependson='ADNI-volumes-prep', fig.width=7, fig.height=7>>=
names(means.complete)[8] = "FreeSurfer"
melted=melt(means.complete, measure.vars=c("FreeSurfer", "FSL", "MAPER","MAGeT"), 
            variable.name = "method", value.name = "volume")

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% ADNI1 X-Validation Dataset Demographics Table
%
<<ADNI-SNT-Xval-demographics, echo=F, results="asis",cache=FALSE>>=
adni_69_RIDs = c(295,413,619,685,729,782,938,954,1018,1155,907,981,222,324,448,
                 546,553,572,602,610,814,1341,675,681,731,1130,41,68,70,101,
                 249,293,316,414,698,1206,1222,1339,751,1030,3,5,10,16,22,53,
                 168,183,241,1205,328,1095,991,213,159,337,343,642,753,1109,
                 1217,29,56,1188,1293,149,382,431,1202)

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-SNT-XVAL: show variability of MAGeT over parameters
\begin{figure}
<<ADN1-SNT-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep', fig.width=5, fig.width=5>>=
ants = subset(snt_xval_data, 
                     reg_method=="ANTS" & 
                     method.mb=="Majority Vote" &
                    templates.mb %in% seq(1,20,2))

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% ADNI1 X-Validation Dataset Demographics Table
%
<<tab:ADNI-xval-demographics, echo=F, dependson='ADNI1-xval-load', results="asis",cache=FALSE>>=
tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + CDRSB + ADAS13 +
               MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
               jens_atlases$RID, method = "reverse", test=FALSE, overall=TRUE)
caption = 
  "\\textbf{ADNI-1 cross-validation subset demographics.} 
  CN     - Cognitively Normal. 
  LMCI   - Late-onset Mild Cognitive Impairment. 
  AD     - Alzheimer's Disease.  
  Hisp   - Hispanic. 
  CDR-SB - Clinical Dementia Rating-Sum of Boxes.  
  ADAS   - Alzheimer's Disease Assessment Scale.  
  MMSE   - Mini-Mental State Examination."
latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
      caption=caption, caption.loc = c("top"),
      label="tab:ADNI1-xvalidation-demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Subfield Validation Template library demographics (healthy controls)
%                                     
<<tab:WAval-healthy-demographics, echo=F, results="asis",cache=FALSE>>=
master = read.csv('data/DT_study.csv')
tab <- summary( Diagnosis ~ Age + Sex + Education + Handedness, data = master,
    method="reverse", na.rm = TRUE, test=FALSE)
caption = 
  "\\textbf{Demographics for the hippocampal subfield cross-validation healthy
control subject sample used in the template library (excluding the Winterburn
atlas subjects).} Education is shown in years."
latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
      caption=caption, caption.loc = c("top"),
      title="tab:WAval-healthy-demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First Episode SZ Demographics Table
%
<<tab:FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
caption = 
  "\\textbf{Schizophrenia First Episode Patient Demographics.}
  ambi - ambidextrous. 
  SES  - Socioeconomic Status score. 
  FSIQ - Full Scale IQ.
"
fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
fep_sz_demographics = subset(fep_sz_demographics, Anon <= 81)
tab <- summary(DX ~ Age + Gender + Handedness + Education + SES + FSIQ,
               data=fep_sz_demographics, method = "reverse", test = FALSE)
latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
      caption=caption, caption.loc = c("top"),
      title="tab:SZFEP-Demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                             
% ADNI1 Complete 1Yr Dataset Demographics Table 
%                                                                             
<<tab:ADNI1-Dataset-Demographics, echo=F, results="asis",cache=FALSE>>=
# RID < 2000 ensures only ADNI1 patients
yr1 = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
yr1$VISCODE <- factor(yr1$Visit, levels = c(1,2,3,4), labels=c("bl","m03","m06","m12"))
yr1 = subset(yr1, !grepl("Scaled_2", Description))
yr1 = yr1[,c("RID","VISCODE")]
yr1 = 
adnimerge.yr1 = merge(adnimerge, yr1) 
tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + CDRSB + ADAS13 +
               MMSE, data = adnimerge.yr1, 
               #subset = RID %in% yr1$RID & VISCODE %in% c('bl','m06','m12'),   
               method = "reverse", test=FALSE, overall=TRUE)
caption = 
  "\\textbf{ADNI1 1.5T Complete 1Yr dataset demographics.} 
  CN     - Cognitively Normal. 
  LMCI   - Late-onset Mild Cognitive Impairment. 
  AD     - Alzheimer's Disease.  
  Hisp   - Hispanic. 
  CDR-SB - Clinical Dementia Rating-Sum of Boxes.  
  ADAS   - Alzheimer's Disease Assessment Scale.  
  MMSE   - Mini-Mental State Examination."
latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
    caption=caption, caption.loc = c("top"),
    title="tab:ADNI1-Dataset-Demographics",
    label="tab:ADNI1-Dataset-Demographics")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: show variability of MAGeT over parameters
\begin{figure}
<<fig:ADN1-Xval-variability,cache=TRUE,dependson='ADNI1-Xval-prep',fig.width=5,fig.height=5>>=
ants = subset(jens_xval_data, 
                     reg_method=="ANTS" & 
                     method.mb=="Majority Vote" &
                    templates.mb %in% seq(1,20,2))

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%                                                                     %%%%%
%%%%%    RESULTS: Experiment 2: Hippocampal Subfield Cross-Validation     %%%%%
%%%%%                                                                     %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment 2 Results: Winterburn Atlases Cross-Validation}
<<WAval-prep, cache=T>>=
wa.xval.vols = na.omit(read.csv('data/WAval.csv'))
wa.xval.casted = dcast(melt(wa.xval.vols, id.vars=c("file","version")), file + variable ~ version)
wa.xval.casted = rename(wa.xval.casted, c("variable"="region"))
levels(wa.xval.casted$region) <-list("CA1"="X1", "CA1"="X101", 
                            "Subiculum"="X2","Subiculum"="X102", 
                            "CA4/DG"="X4","CA4/DG"="X104", 
                            "CA2/CA3"="X5","CA2/CA3"="X105", 
                            "SR/SL/SM"="X6","SR/SL/SM"="X106")

wa.xval.error = ddply(wa.xval.casted, c("file", "region"), function(df) {
  with(df, 
  data.frame(
    resampled = (gold_0.9mm - gold_0.3mm) / gold_0.3mm * 100, 
    mb_bravo  = (mb_bravo   - gold_0.3mm) / gold_0.3mm * 100, 
    mb_0.9mm  = (mb_0.9mm   - gold_0.3mm) / gold_0.3mm * 100))})

wa.s = ddply(wa.xval.error, c('region'), function(df) {
  with(df,data.frame(
    m_res = mean(resampled,na.rm=T), sd_res = sd(resampled,na.rm=T),
    m_bravo = mean(mb_bravo,na.rm=T), sd_bravo = sd(mb_bravo,na.rm=T),
    m_mb_res = mean(mb_0.9mm,na.rm=T), sd_mb_res = sd(mb_0.9mm,na.rm=T)))})
@

In this experiment the \mb{} segmentation of hippocampal subfields is assessed
through a leave-one-out validation in which resampled images ($0.9mm^3$
voxels) of each Winterburn atlas subject are segmented using the remaining
high-resolution Winterburn atlas subjects' images.

Figure \ref{fig:WAval-vol-boxplot}) shows the mean relative percent error in volume, 
with respect to the full-resolution Winterburn atlas, of each hippocampal subfield
of the resampled Winterburn labels, and of the \mb{} segmentations of the resampled
Winterburn atlas images and T1 BRAVO Winterburn subject acquisitions.  A
paired-samples t-test was conducted to compare the distribution of relative
percent error in the \mb{} segmentation volumes and the resampled segmentation
volumes. There was no significant difference found for the CA1, Subiculum, and
CA2/CA3 subregions. The CA4/DG subregion shows significantly lower percent error 
  ($M_{res}=$\Sexpr{wa.s$m_res[wa.s$region=='CA4/DG']},
  $SD_{res}=$\Sexpr{wa.s$sd_res[wa.s$region=='CA4/DG']};
  $M_{mb-bravo}=$\Sexpr{wa.s$m_bravo[wa.s$region=='CA4/DG']},    
  $SD_{mb-bravo}=$\Sexpr{wa.s$sd_bravo[wa.s$region=='CA4/DG']};    
  $M_{mb-res}=$\Sexpr{wa.s$m_mb_res[wa.s$region=='CA4/DG']},    
  $SD_{mb-res}=$\Sexpr{wa.s$sd_mb_res[wa.s$region=='CA4/DG']};    
  $p < 0.001$), and the SR/SL/SM shows significantly different percent error in the opposite
direction 
  ($M_{res}=$\Sexpr{wa.s$m_res[wa.s$region=='SR/SL/SM']},
  $SD_{res}=$\Sexpr{wa.s$sd_res[wa.s$region=='SR/SL/SM']};
  $M_{mb-bravo}=$\Sexpr{wa.s$m_bravo[wa.s$region=='SR/SL/SM']},    
  $SD_{mb-bravo}=$\Sexpr{wa.s$sd_bravo[wa.s$region=='SR/SL/SM']};    
  $M_{mb-res}=$\Sexpr{wa.s$m_mb_res[wa.s$region=='SR/SL/SM']},    
  $SD_{mb-res}=$\Sexpr{wa.s$sd_mb_res[wa.s$region=='SR/SL/SM']};    
  $p < 0.001$). Bland-Altman plots comparing \mb{} to subsampled segmentation
volume reveals a proportional bias towards underestimating volumes in all
subregions, and a noticeable fixed bias in the CA4/DG and SR/SL/SM subregions.

Figure \ref{fig:subfield-montage} shows slices subfield segmentations for a
single subject for qualitative inspection.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Winterburn-XVAL: compare resampling
\begin{figure}[h!]
<<fig:WAval-vol-boxplot,cache=T,fig.width=6,fig.height=5>>=

@@

ggplot(wa.xval.error.melted, aes(y=percenterr,x=measure,colour=measure)) +
  facet_grid(.~region) + 
  geom_hline(aes(yintercept=0), colour="grey", size=1) + 
  geom_boxplot() + 
  geom_point(group=1, position="dodge") + 
  labs(y="Percent error in volume", x="Subfield") +
  theme(axis.ticks = element_blank(), axis.text.x = element_blank()) + 
  scale_colour_discrete("") + 
  theme(legend.direction = "vertical", legend.position = "bottom")
@
  \caption{{\bf Percent error in computed unilateral hippocampal subfield
segmentation volume relative to manual segmentation volume.} 
  Percent error is measured against the volumes of the unmodified Winterburn
  atlas subfield segmentations.
  {\bf Subsampled atlas labels} volumes of the manual segmentations of the Winterburn
  atlases after resampling to $0.9mm^3$ voxels.  
  {\bf MAGeT on subsampled atlas images} volumes are \mb{} segmentations of the Winterburn atlas
  images after resampling to $0.9mm^3$ voxels.
  {\bf MAGeT on T1 BRAVO} volumes are \mb{} segmentations of T1 BRAVO images
  ($0.9mm^3$ voxels) acquired separately of four of the five Winterburn atlas
  subjects.\label{fig:WAval-vol-boxplot}}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Winterburn-XVAL: bland-altman 
\begin{figure}[h!]
<<fig:WAval-bland,cache=T,fig.width=7,fig.height=6>>=
wa.xval.volumes = melt(wa.xval.casted, id.vars=c('file', 'region', 'gold_0.3mm', 'gold_0.9mm'))
wa.xval.volumes = rename(wa.xval.volumes, c("variable"="measure", "value"="volume"))
levels(wa.xval.volumes$measure) <- list("Subsampled atlas labels"="gold_0.9mm",
                               "MAGeT on subsampled atlas images" = "mb_0.9mm",
                               "MAGeT on T1 BRAVO images"="mb_bravo")

@@

ggplot(wa.xval.volumes, aes(x=mean, y=diff, colour=measure)) + 
  facet_wrap(~region, ncol=3, scales='free') + 
  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
  geom_point(size=1) +  
  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
             data=wa.xval.bland.limits, size=0.3, alpha=0.5) + 
  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", 
             data = wa.xval.bland.limits, hjust=1.1,vjust=-0.8, size=2.5) + 
  xlab(expression('Mean of resampled and \\mb subregion volume ' (mm^3))) + 
  ylab(expression('Difference between manual and \\mb subregion volume ' (mm^3))) + 
  scale_color_discrete(name="") + 
  theme(legend.direction = "horizontal", legend.position = "bottom", 
        axis.text.y = element_text(angle = 90, hjust=0.5), 
        axis.text = element_text(size = 8))

@
  \caption{{\bf Bland-Altman plots comparing unilateral \mb{} subfield
segmentation volume with resampled subfield segmentation volume.} The overall
mean difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by
dashed horizontal lines. Linear fit lines are shown for each target image type.
Note, points below the mean difference indicate overestimation of the volume
with respect to the resampled volume, and vice versa. 
  \label{fig:WAval-bland}}
\end{figure}

@@

\begin{figure}
<<fig:FEP-volumes,cache=FALSE,dependson="setup",fig.width=5,fig.height=5>>=
SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")
mean_sz_vols = ddply(SZ_volumes, c('Subject', 'Method'), function (df) {
  data.frame(Volume = mean(df$Volume))})

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1 1yr Complete Dataset Package Totals & Failures
%                        
<<tab:ADNI-seg-package-totals, echo=F, results="asis",cache=FALSE>>=
package_totals$Total <- NULL
latex(package_totals, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
    caption="\\textbf{Number of segmented images and quality control failures of 
             \\adnidataset{} dataset by method.}", 
    caption.loc = c("top"), rowname=NULL, 
    label="tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h]
<<fig:ADNI-volumes-boxplot,dependson='ADNI-volumes-prep',cache=T,fig.width=7,fig.height=4>>=
melted = melt(means.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT"),
              variable.name="Method", value.name="Volume")
correlations = ddply(melted, c("DX","Method"), function (df) {
  data.frame(
    pearson = cor(df$SNT, df$Volume)
  )
})
m2 = melt(means.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT","SNT"), variable.name="Method")
levels(m2$Method) = c("FS"    = "FreeSurfer", 
                      "FSL"   = "FSL", 
                      "MAPER" = "MAPER", 
                      "MAGeT" = "MAGeT", 
                      "SNT"   = "SNT")
m2$DX = factor(m2$DX, levels(m2$DX)[c(2,3,1)])

@@

\begin{figure}[h]
<<fig:ADNI-volumes-plot,dependson='ADNI-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
melted=melt(means.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
            variable.name = "Method", value.name = "volume")

@@

\begin{figure}
<<fig:ADNI-Bland-Altman, cache=TRUE, dependson='ADNI-volumes-prep', fig.width=7, fig.height=7>>=
names(means.complete)[8] = "FreeSurfer"
melted=melt(means.complete, measure.vars=c("FreeSurfer", "FSL", "MAPER","MAGeT"), 
            variable.name = "method", value.name = "volume")

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% ADNI1 X-Validation Dataset Demographics Table
%
<<tab:ADNI-SNT-Xval-demographics, echo=F, results="asis",cache=FALSE>>=
adni_69_RIDs = c(295,413,619,685,729,782,938,954,1018,1155,907,981,222,324,448,
                 546,553,572,602,610,814,1341,675,681,731,1130,41,68,70,101,
                 249,293,316,414,698,1206,1222,1339,751,1030,3,5,10,16,22,53,
                 168,183,241,1205,328,1095,991,213,159,337,343,642,753,1109,
                 1217,29,56,1188,1293,149,382,431,1202)

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-SNT-XVAL: show variability of MAGeT over parameters
\begin{figure}
<<fig:ADN1-SNT-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep', fig.width=5, fig.width=5>>=
ants = subset(snt_xval_data, 
                     reg_method=="ANTS" & 
                     method.mb=="Majority Vote" &
                    templates.mb %in% seq(1,20,2))
